Reproduce anaylses of Beninca et al (2008)

### 2015-10-20 code/report.Rmd
update

This is a reproduction of the anaylses presented in the paper *Chaos in a long-term experiment with a plankton community*, by Elisa Benincà and others ([the paper on the Nature website](http://www.nature.com/nature/journal/v451/n7180/abs/nature06512.html)). Details of the methods are in [the Supplement to the Nature paper](http://www.nature.com/nature/journal/v451/n7180/extref/nature06512-s1.pdf).

@@

The code and data for the reproduction are [here](https://github.com/opetchey/RREEBES/tree/Beninca_development) on github.

@@

The datasets for this reproduction are read from this github repository:
```{r}
repo <- "https://raw.githubusercontent.com/opetchey/ReScience-submission/beninca_dev"
```

# First get the raw data into R and tidy it.

@@

First quick go:
```{r}
ggplot(all.data, aes(x=Day.number, y=value)) + geom_line() +
  facet_wrap(~variable, scales="free_y")
```

Now we add a column that gives the variable types, same as in figure 1b through 1g.
First make a lookup table giving species type:
```{r}
tt <- data.frame(variable=unique(all.data$variable),
                 Type=c("Cyclopoids", "Herbivore", "Herbivore", "Herbivore",
                        "Phytoplankton",  "Phytoplankton", "Phytoplankton",
                        "Detritivore", "Detritivore", "Bacteria", "Nutrient", "Nutrient"))
#tt
```

@@

```{r}
all.data1 <- na.omit(all.data)

#group_by(all.data1, variable) %>% summarise(out=min(Day.number))

mt <- plyr::dlply(all.data1,
                  "variable",
                  function(xx) pracma::interp1(x=xx$Day.number,
                                               y=xx$value,
                                               xi=xout,
                                               method="cubic"))
## Aside: the duplicated date that was previously fixed was only discovered by a warning message
## given by the pracma::interp1 function!!!

mt <- as.data.frame(mt)
mt <- cbind(Day.number=xout, mt)
mt <- gather(mt, variable, value, 2:13)
#ggplot(mt, aes(x=Day.number, y=value)) + facet_wrap(~variable, scales="free") + geom_line()
```

@@

from.steve <- read.csv(text=getURL(paste0(repo,"/data/direct_from_Steve/interp_short_allsystem_newnames.csv")), header=T)

@@

Looks OK, but suggests / shows that that data in the ELE supplement were standardised after removal of the zeros, whereas we don't do any zero removal (this is the same as in Ellner's code.)

@@

Get p-vals too:
```{r}
cor.pvals <- matrix(NA, length(final_wide[1,]), length(final_wide[1,]))
for(i in 1:length(final_wide[1,]))
  for(j in 1:length(final_wide[1,])) {
    if(sum(!is.na(final_wide[,i]) & !is.na(final_wide[,j]))>10)
    cor.pvals[i,j] <- cor.test(final_wide[,i], final_wide[,j])$p.value
}
    #plot(final_wide[,2], final_wide[,3])
```

@@

```{r}
# read script lines from website
script <- getURL(paste0(repo,"/report/functions/indirect_method_functions.R"), ssl.verifypeer = FALSE)

@@

Get the correlations and do some house keeping:

```{r}
cors <- matrix(NA, 12, 12)
dimnames(cors) = list(dist=1:12,
                        species=names(x))
for(i in 1:12) {
  for(j in 1:12) {
    cors[j,i] <- cor(Z1[1:nn,i], preds[j, i, ], use="complete.obs")^2
}}

@@

And plot our version of figure 2:

```{r}
levels(cors.long$Variable)
ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
  geom_point() +
  facet_wrap(~Variable, scales="free_y", ncol=3 )

cors.long$Variable <- factor(cors.long$Variable, sort(levels(cors.long$Variable)))
ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
  geom_point() +
  facet_wrap(~Variable, ncol=3 )

cors.long$Variable <- factor(cors.long$Variable, c("Cyclopoids", "Rotifers", "Calanoid.copepods", "Picophytoplankton",
                             "Nanophytoplankton", "Filamentous.diatoms", "Soluble.reactive.phosphorus",
                             "Total.dissolved.inorganic.nitrogen", "Bacteria", "Ostracods", "Harpacticoids",
                             "Protozoa"))
ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
  geom_point(col="red", size=3) +
  facet_wrap(~Variable, ncol=3 )

```{r}
original_preds <- read.csv(text=getURL(paste0(repo,"/data/original_rsquared.csv")), skip=0, header=T)
opw <- gather(original_preds, Species, r_squared, 2:25)
opw <- separate(opw, Species, c("Species", "Model"))

@@

First pass working. Need to double check everything. Quite different patterns from in figure 2 of the nature paper. There is relatively little information in the paper or supplement about how figure 2 data was produced, so difficult to pin down the difference without asking authors.

>>>>

This document contains various reproductions of the anaylses presented in the paper *Chaos in a long-term experiment with a plankton community*, by Elisa Benincà and others ([the paper on the Nature website](http://www.nature.com/nature/journal/v451/n7180/abs/nature06512.html)). Details of the methods are in [the Supplement to the Nature paper](http://www.nature.com/nature/journal/v451/n7180/extref/nature06512-s1.pdf).

@@

*Update* The code and data for the reproduction are [here](https://github.com/opetchey/RREEBES/tree/Beninca_development) on github.

@@

The datasets are read from this github repository:
```{r}
repo <- "https://raw.githubusercontent.com/opetchey/ReScience-submission/beninca_dev"
```

# First get the raw data into R, and clean and tidy it.

@@

First add a column that gives the variable types, same as in figure 1b through 1g. Make a table giving species type:
```{r}
tt <- data.frame(variable=unique(all.data$variable),
                 Type=c("Cyclopoids", "Herbivore", "Herbivore", "Herbivore",
                        "Phytoplankton",  "Phytoplankton", "Phytoplankton",
                        "Detritivore", "Detritivore", "Bacteria", "Nutrient", "Nutrient"))
#tt
```

@@

Perform the interpolation:
```{r}
all.data1 <- na.omit(all.data)
mt <- plyr::dlply(all.data1,
                  "variable",
                  function(xx) pracma::interp1(x=xx$Day.number,
                                               y=xx$value,
                                               xi=xout,
                                               method="cubic"))
## Aside: the duplicated date that was previously fixed was only discovered by a warning message
## given by the pracma::interp1 function!

mt <- as.data.frame(mt)
mt <- cbind(Day.number=xout, mt)
mt <- gather(mt, variable, value, 2:13)
```

@@

from.steve <- read.csv(text=getURL(paste0(repo,"/data/interp_short_allsystem_newnames.csv")), header=T)

@@

Looks OK, but suggests / shows that that data in the ELE supplement were standardised after removal of the zeros.

@@

Get p-vals:
```{r}
cor.pvals <- matrix(NA, length(final_wide[1,]), length(final_wide[1,]))
for(i in 1:length(final_wide[1,]))
  for(j in 1:length(final_wide[1,])) {
    if(sum(!is.na(final_wide[,i]) & !is.na(final_wide[,j]))>10)
    cor.pvals[i,j] <- cor.test(final_wide[,i], final_wide[,j])$p.value
}
    #plot(final_wide[,2], final_wide[,3])
```

@@

```{r}
# read script lines from website
script <- getURL(paste0(repo,"/code/indirect_method_functions.R"), ssl.verifypeer = FALSE)

@@

Get the correlations and do some house keeping:
```{r}
cors <- matrix(NA, 12, 12)
dimnames(cors) = list(dist=1:12,
                        species=names(x))
for(i in 1:12) {
  for(j in 1:12) {
    cors[j,i] <- cor(Z1[1:nn,i], preds[j, i, ], use="complete.obs")^2
}}

@@

And plot our version of figure 2:
```{r}

cors.long$Variable <- factor(cors.long$Variable, c("Cyclopoids", "Rotifers", "Calanoid.copepods", "Picophytoplankton",
                             "Nanophytoplankton", "Filamentous.diatoms", "Soluble.reactive.phosphorus",
                             "Total.dissolved.inorganic.nitrogen", "Bacteria", "Ostracods", "Harpacticoids",
                             "Protozoa"))
# ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
#   geom_point(col="red", size=3) +
#   facet_wrap(~Variable, ncol=3 )

## read in the data of the original article that came direct from Elisa
original_preds <- read.csv(text=getURL(paste0(repo,"/data/original_rsquared.csv")), skip=0, header=T)
opw <- gather(original_preds, Species, r_squared, 2:25)
opw <- separate(opw, Species, c("Species", "Model"))

@@

Quite different patterns from in figure 2 of the nature paper. There is relatively little information in the paper or supplement about how figure 2 data was produced, so difficult to pin down the difference without asking authors.
